---
output:
  md_document:
    variant: gfm+tex_math_dollars
---

# MCPower validation — OLS solver (data → results)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Run from mcpower/validation/ (rmarkdown::render sets wd to this file's dir).
source("common.R")     # build_model, fit_sota, fit_mcpower, solving_rows, hash_check, ...
source("tolerances.R") # SOLVING_TOL — the central B<->C gate
source("formulas.R")   # CASES
OLS_CASES <- Filter(function(c) c$family == "ols", CASES)
DATA_DIR  <- "data"

png_type <- if (capabilities("cairo")) "cairo" else "Xlib"

family_long <- function(case) "ordinary least squares (continuous outcome)"

model_equation <- function(case) {
  e  <- parse_effects(case$effects)
  rhs <- paste(sprintf("%.2f·%s", e, names(e)), collapse = " + ")
  sprintf("y = %s + noise", rhs)
}
```

# What this report shows

MCPower generates a dataset from a formula, then **fits it back** to recover the
coefficients. This report checks the *fitting* half: given one fixed dataset, does
MCPower's own solver return the same numbers a trusted, independent R fit returns?

For each formula we take the exact bytes MCPower saved (the same design matrix and
outcome, byte-for-byte), fit them two ways —

- **R (the reference):** `stats::lm` through the origin on the design matrix, the
  textbook least-squares solution.
- **MCPower:** the engine's own `load_data()` path, which runs the *same* solver the
  power simulation uses.

— and compare the fitted coefficients, the test statistics, and the decision
threshold. Because both fit the identical bytes, sampling noise cancels: any
difference is a solver discrepancy, and the allowed differences are tiny.

# How the check works

1. **Provenance.** Re-generate the dataset from the committed seed and confirm its
   content hash matches the saved file — proof the bytes haven't drifted.
2. **B (R) vs C (MCPower)** on those bytes: coefficients, the marginal *t*
   statistics, and the critical value.
3. **C vs A** (a readable sanity overlay): MCPower's recovered coefficients next to
   the formula's true values.

# The thresholds

| Quantity | Allowed difference | Why |
|---|---|---|
| Coefficient (β) | `r SOLVING_TOL$estimate_rel_ols` relative | exact normal equations; observed worst agreement ≈1.2e-12 |
| Statistic (*t*) | `r SOLVING_TOL$estimate_rel_ols` relative | derived from the same β and variance |
| Critical value | `r SOLVING_TOL$crit_abs` absolute | `qt` quantile agreement ~1e-15 |

These are the B↔C gates from `tolerances.R`. OLS is closed-form, so it gets the
**tight** `estimate_rel_ols` band — set just above the measured worst agreement
(≈1.2e-12) so a real solver regression would trip it. The iterative GLM/MLE fits
use the looser `estimate_rel_iter` band in their own reports; the per-formula
tables below show OLS's actual margin.

```{r compute}
RESULTS <- lapply(OLS_CASES, function(case) {
  saved <- readRDS(file.path(DATA_DIR, paste0(case$label, ".rds")))
  hc    <- hash_check(case)                    # provenance (common.R)
  sota  <- sota_side(case, saved)
  mc    <- mcpower_side(case, saved)
  sr    <- compare_sides(case, sota, mc, saved)
  list(case = case, saved = saved, hash_ok = hc$ok, sota = sota,
       table = sr$table, converged = sr$converged)
})
names(RESULTS) <- vapply(OLS_CASES, function(c) c$label, character(1))
```

```{r summary}
summ <- data.frame(
  Formula  = vapply(RESULTS, function(r) r$case$formula, character(1)),
  n        = vapply(RESULTS, function(r) as.integer(r$case$n), integer(1)),
  `B↔C`    = vapply(RESULTS, function(r)
               if (all(r$table$Verdict == "PASS")) "PASS" else "FAIL", character(1)),
  Converged= vapply(RESULTS, function(r) if (r$converged) "yes" else "NO", character(1)),
  Reproduces = vapply(RESULTS, function(r) if (r$hash_ok) "yes" else "NO", character(1)),
  row.names = NULL, check.names = FALSE)
knitr::kable(summ)
```

```{r overall, results='asis'}
all_ok <- all(vapply(RESULTS, function(r)
  all(r$table$Verdict == "PASS") && r$converged && r$hash_ok, logical(1)))
cat(if (all_ok)
  "> **All OLS formulas pass:** MCPower's solver matches R to machine precision on every saved dataset, and every file reproduces from its seed.\n"
  else "> **Some formulas need review** — see the per-formula sections below.\n")
stopifnot(all_ok)   # G5: render fails on any FAIL, no silent green HTML
```

```{r agreement-plot, results='asis'}
# Figures land beside this report's md in the published web tree; the md is
# rendered with output_dir = web/documentation/validation, so a relative
# "figures/<x>.png" src resolves at the served page (mirrors data_generation.rmd).
figdir_src <- "figures"
figdir_out <- file.path("..", "web", "documentation", "validation", figdir_src)
dir.create(figdir_out, recursive = TRUE, showWarnings = FALSE)
fpath <- file.path(figdir_out, "ols_agreement.png")
grDevices::png(fpath, width = 780, height = 280, res = 96, type = png_type)
plot_solver_agreement(RESULTS, SOLVING_TOL$estimate_rel_ols,
                      sprintf("%g rel", SOLVING_TOL$estimate_rel_ols))
grDevices::dev.off()
cat(sprintf("![B↔C relative agreement of every coefficient and statistic across all OLS formulas, against the tolerance gate.](%s)\n\n",
            file.path(figdir_src, "ols_agreement.png")))
cat("*Each point is one fitted quantity on one saved dataset; the red line is the gate. Every β and statistic agrees with R orders of magnitude inside tolerance — exact (bit-identical) matches pile up at the left floor.*\n\n")
```

```{r golden-freeze}
# Freeze R-SOTA's B side for every case (campaign is the sole golden producer).
invisible(lapply(RESULTS, function(r) write_golden(r$case, r$saved, r$sota)))
```

```{r per-formula, results='asis'}
for (r in RESULTS) {
  cat(sprintf("\n## %s\n\n", model_equation(r$case)))
  cat(sprintf("R formula `%s` &middot; %s &middot; n=%d, seed=%d. ",
              r$case$formula, family_long(r$case), r$case$n, r$case$seed))
  cat(sprintf("File reproduces from seed: **%s**.\n\n",
              if (r$hash_ok) "yes" else "NO"))
  cat(knitr::kable(compact_solver_table(r$table), digits = 6, format = "pipe"), sep = "\n")
  cat("\n\n")
}
```

```{r provenance}
prov <- data.frame(
  item  = c("generated", "R", "mcpower", "OLS solving bands (beta/stat rel · crit abs)"),
  value = c(as.character(Sys.Date()), R.version.string,
            as.character(utils::packageVersion("mcpower")),
            sprintf("%g rel / %g rel / %g abs", SOLVING_TOL$estimate_rel_ols,
                    SOLVING_TOL$estimate_rel_ols, SOLVING_TOL$crit_abs)),
  check.names = FALSE)
knitr::kable(prov)
```

Reproduce: `Rscript mcpower/validation/data_generation.r` then
`rmarkdown::render("mcpower/validation/validation_OLS_solving.rmd", output_dir = "mcpower/web/documentation/validation")`.
