---
output:
  md_document:
    variant: gfm+tex_math_dollars
---

# MCPower validation — GLM solver (data → results)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Run from mcpower/validation/ (rmarkdown::render sets wd to this file's dir).
source("common.R")     # build_model, fit_sota, fit_mcpower, solving_rows, hash_check, ...
source("tolerances.R") # SOLVING_TOL — the central B<->C gate
source("formulas.R")   # CASES
GLM_CASES <- Filter(function(c) c$family == "logit", CASES)
DATA_DIR  <- "data"

png_type <- if (capabilities("cairo")) "cairo" else "Xlib"

family_long <- function(case) "logistic regression (binary outcome)"

model_equation <- function(case) {
  e <- parse_effects(case$effects)
  parts <- vapply(seq_along(e), function(i)
    sprintf("%s %g·%s", if (e[i] < 0) "−" else "+", abs(e[i]), names(e)[i]), character(1))
  body <- paste(parts, collapse = " ")
  sprintf("log-odds(y = 1) = logit(%.2f) %s", case$baseline_probability, body)
}
```

# What this report shows

MCPower generates a binary dataset from a formula, then **fits it back** to recover
the coefficients. This report checks the *fitting* half: given one fixed dataset,
does MCPower's own solver return the same numbers a trusted, independent R fit
returns?

For each formula we take the exact bytes MCPower saved (the same design matrix and
0/1 outcome, byte-for-byte), fit them two ways —

- **R (the reference):** `stats::glm(family = binomial())` through the origin on the
  design matrix, the textbook maximum-likelihood logistic fit.
- **MCPower:** the engine's own `load_data()` path, which runs the *same* iterative
  IRLS solver the power simulation uses.

— and compare the fitted coefficients, the test statistics, and the decision
threshold. Because both fit the identical bytes, sampling noise cancels: any
difference is a solver discrepancy, and the allowed differences are small.

# How the check works

1. **Provenance.** Re-generate the dataset from the committed seed and confirm its
   content hash matches the saved file — proof the bytes haven't drifted.
2. **B (R) vs C (MCPower)** on those bytes: coefficients, the Wald *z*
   statistics (β̂/se), and the critical value `qnorm(1 − α/2)` ≈ 1.96.
3. **C vs A** (a readable sanity overlay): MCPower's recovered coefficients next to
   the formula's true values (on the log-odds scale).

# The thresholds

| Quantity | Allowed difference | Why |
|---|---|---|
| Coefficient (β) | `r SOLVING_TOL$estimate_rel_iter` relative | iterative IRLS fit — agrees with R well inside this band |
| Statistic (*z*) | `r SOLVING_TOL$estimate_rel_iter` relative | derived from the same β and variance |
| Critical value | `r SOLVING_TOL$crit_abs_iter` absolute | engine's own normal/χ² quantile vs R's `qnorm` ≈1.6e-9 |

These are the B↔C gates from `tolerances.R`. GLM is an **iterative IRLS** fit, so its
agreement with R is looser than OLS's closed-form solution: it gets the
`estimate_rel_iter` band (typically ~1e-6 to 1e-9 for well-converged IRLS), not the
tight OLS `estimate_rel_ols`. The per-formula tables below show each formula's actual
margin.

```{r compute}
RESULTS <- lapply(GLM_CASES, function(case) {
  saved <- readRDS(file.path(DATA_DIR, paste0(case$label, ".rds")))
  hc    <- hash_check(case)                    # provenance (common.R)
  sota  <- sota_side(case, saved)
  mc    <- mcpower_side(case, saved)
  sr    <- compare_sides(case, sota, mc, saved)
  list(case = case, saved = saved, hash_ok = hc$ok, sota = sota,
       table = sr$table, converged = sr$converged)
})
names(RESULTS) <- vapply(GLM_CASES, function(c) c$label, character(1))
```

```{r summary}
summ <- data.frame(
  Formula  = vapply(RESULTS, function(r) r$case$formula, character(1)),
  n        = vapply(RESULTS, function(r) as.integer(r$case$n), integer(1)),
  `B↔C`    = vapply(RESULTS, function(r)
               if (all(r$table$Verdict == "PASS")) "PASS" else "FAIL", character(1)),
  Converged= vapply(RESULTS, function(r) if (r$converged) "yes" else "NO", character(1)),
  Reproduces = vapply(RESULTS, function(r) if (r$hash_ok) "yes" else "NO", character(1)),
  row.names = NULL, check.names = FALSE)
knitr::kable(summ)
```

```{r overall, results='asis'}
all_ok <- all(vapply(RESULTS, function(r)
  all(r$table$Verdict == "PASS") && r$converged && r$hash_ok, logical(1)))
cat(if (all_ok)
  "> **All GLM formulas pass:** MCPower's IRLS solver matches R's `glm` well inside the iterative band on every saved dataset, and every file reproduces from its seed.\n"
  else "> **Some formulas need review** — see the per-formula sections below.\n")
stopifnot(all_ok)   # G5: render fails on any FAIL, no silent green HTML
```

```{r golden-freeze}
# Freeze R-SOTA's B side for every case (campaign is the sole golden producer).
invisible(lapply(RESULTS, function(r) write_golden(r$case, r$saved, r$sota)))
```

```{r per-formula, results='asis'}
for (r in RESULTS) {
  cat(sprintf("\n## %s\n\n", model_equation(r$case)))
  cat(sprintf("R formula `%s` &middot; %s &middot; n=%d, seed=%d. ",
              r$case$formula, family_long(r$case), r$case$n, r$case$seed))
  cat(sprintf("File reproduces from seed: **%s**.\n\n",
              if (r$hash_ok) "yes" else "NO"))
  cat(knitr::kable(r$table, digits = 6, format = "pipe"), sep = "\n")
  cat("\n\n")
}
```

```{r provenance}
prov <- data.frame(
  item  = c("generated", "R", "mcpower", "GLM solving bands (beta/stat rel · crit abs)"),
  value = c(as.character(Sys.Date()), R.version.string,
            as.character(utils::packageVersion("mcpower")),
            sprintf("%g rel / %g rel / %g abs", SOLVING_TOL$estimate_rel_iter,
                    SOLVING_TOL$estimate_rel_iter, SOLVING_TOL$crit_abs_iter)),
  check.names = FALSE)
knitr::kable(prov)
```

Reproduce: `Rscript mcpower/validation/data_generation.r` then
`rmarkdown::render("mcpower/validation/validation_GLM_solving.rmd", output_dir = "mcpower/documentation/validation")`.
