---
output:
  md_document:
    variant: gfm+tex_math_dollars
---

# MCPower validation — get_effects_from_data round-trip

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Run from mcpower/validation/ (rmarkdown::render sets wd to this file's dir).
source("common.R")     # build_model, continuous_names, parse_effects
source("tolerances.R") # GETEFFECTS_TOL — the central round-trip gate
source("formulas.R")   # CASES

# Campaign parameters (not tolerances — see tolerances.R for the gate).
GE_N     <- 4000L      # per-draw sample size (LME scales n_clusters to ~match)
GE_K     <- 20L        # draws averaged (single-draw noise ~1/sqrt(n) averaged down)
GE_SEED0 <- 2137L      # first draw seed; draws use GE_SEED0 + 0..K-1

# GE_CASES, scale_case, recover_once, expected_recovery are defined in
# formulas.R (GE_CASES) and common.R (the three helpers) — sourced above.

# K-draw round-trip for one case: mean recovered vs expected, with the per-term
# verdict against GETEFFECTS_TOL$mean_abs.
roundtrip_rows <- function(case) {
  cc    <- scale_case(case, GE_N)
  cont  <- continuous_names(cc)
  draws <- matrix(
    vapply(seq_len(GE_K), function(k) recover_once(cc, GE_SEED0 + k - 1L),
           numeric(length(cont))),
    nrow = length(cont))
  mean_rec <- rowMeans(draws)
  exp_v    <- as.numeric(expected_recovery(cc))
  abs_err  <- abs(mean_rec - exp_v)
  tbl <- data.frame(
    Term            = cont,
    `Specified (s)` = as.numeric(parse_effects(cc$effects)[cont]),
    `Expected`      = exp_v,
    `Mean recovered`= mean_rec,
    `|err|`         = abs_err,
    Verdict         = ifelse(abs_err <= GETEFFECTS_TOL$mean_abs, "PASS", "FAIL"),
    check.names = FALSE, row.names = NULL)
  list(case = cc, table = tbl)
}
```

# What this report validates

`get_effects_from_data` recovers standardized effect sizes from uploaded pilot
data by fitting the model the data came from. This report is the **acceptance gate
for that recovery convention**: for each estimator we *specify* an effect of size
`s`, let MCPower *simulate* a dataset from it, then *recover* the effect from that
dataset — and check the recovery lands where the convention says it should.

The estimator is dispatched off the model family, and the recovery scale differs
by estimator:

- **OLS (continuous outcome)** z-scores the outcome, so it recovers the
  *standardized regression coefficient* — `s / sqrt(Σs² + 1)` for independent
  continuous predictors with the engine's default N(0, 1) residual. The shrinkage
  is the residual variance the standardization divides through; it is small for
  the small/medium effects benchmarked here and is the same approximation the
  shipped OLS recovery has always carried.
- **GLM (logistic)** and **MLE (mixed)** fit the **native** outcome (raw 0/1 for
  logit, the raw response for mixed), so they recover the coefficient **directly**:
  expected recovery is `s` itself. MLE recovery is fixed-effects-only and reads the
  grouping column from the uploaded data.

Scope: **continuous main effects only** — the settled convention. Factor-dummy and
interaction scaling under non-continuous outcomes is a separate, still-open item
and is intentionally not gated here.

# How the check works

1. **Simulate.** Generate a dataset from each case's data-generating process at
   n = `r GE_N` (LME scales the cluster count to match), via the engine's own
   `create_data()` — the same DGP the power simulation uses.
2. **Recover.** Rebuild the raw predictor frame (plus the outcome, plus the
   grouping column for LME), `upload_data()` it, and call
   `get_effects_from_data()`.
3. **Average.** Repeat over `r GE_K` draws (seeds `r GE_SEED0`…`r GE_SEED0 + GE_K - 1L`)
   and take the mean — single-draw noise is ~1/sqrt(n), so averaging isolates the
   convention from sampling scatter (the A↔B harness pattern).
4. **Gate.** The mean recovered effect must land within
   `r GETEFFECTS_TOL$mean_abs` (absolute) of the convention-predicted value.

# The threshold

| Quantity | Allowed difference | Why |
|---|---|---|
| Mean recovered − expected | `r GETEFFECTS_TOL$mean_abs` absolute | K-draw mean vs the convention-predicted value; worst measured margin 0.007 → ~2.8× headroom |

This is the `GETEFFECTS_TOL` gate from `tolerances.R`. A wrong estimator, a sign
flip, a missing/inverted standardization, or a 2× scale would move the mean by
0.1–0.5 — far outside this band.

```{r compute}
RESULTS <- lapply(GE_CASES, roundtrip_rows)
names(RESULTS) <- vapply(RESULTS, function(r) r$case$label, character(1))
```

```{r summary}
fam_long <- c(ols = "OLS (continuous)", logit = "GLM (logistic)", lme = "MLE (mixed)")
summ <- data.frame(
  Case    = vapply(RESULTS, function(r) r$case$label, character(1)),
  Estimator = vapply(RESULTS, function(r) fam_long[[r$case$family]], character(1)),
  Formula = vapply(RESULTS, function(r) r$case$formula, character(1)),
  n       = vapply(RESULTS, function(r) as.integer(r$case$n), integer(1)),
  K       = GE_K,
  Verdict = vapply(RESULTS, function(r)
              if (all(r$table$Verdict == "PASS")) "PASS" else "FAIL", character(1)),
  row.names = NULL, check.names = FALSE)
knitr::kable(summ)
```

```{r overall, results='asis'}
all_ok <- all(vapply(RESULTS, function(r) all(r$table$Verdict == "PASS"), logical(1)))
cat(if (all_ok)
  "> **All estimators round-trip:** every specified continuous main effect is recovered to within the gate of the convention-predicted value — OLS at the shrunk standardized scale, GLM and MLE at the native coefficient.\n"
  else "> **Some cases need review** — see the per-case sections below.\n")
```

```{r per-case, results='asis'}
for (r in RESULTS) {
  cat(sprintf("\n## %s &middot; %s\n\n", r$case$label, fam_long[[r$case$family]]))
  cat(sprintf("R formula `%s` &middot; n=%d &middot; K=%d draws (seeds %d–%d).\n\n",
              r$case$formula, r$case$n, GE_K, GE_SEED0, GE_SEED0 + GE_K - 1L))
  cat(knitr::kable(r$table, digits = 5, format = "pipe"), sep = "\n")
  cat("\n\n")
}
```

```{r provenance}
prov <- data.frame(
  item  = c("generated", "R", "mcpower", "round-trip gate (mean abs)",
            "n / K / seed0"),
  value = c(as.character(Sys.Date()), R.version.string,
            as.character(utils::packageVersion("mcpower")),
            sprintf("%g abs", GETEFFECTS_TOL$mean_abs),
            sprintf("%d / %d / %d", GE_N, GE_K, GE_SEED0)),
  check.names = FALSE)
knitr::kable(prov)
```

Reproduce:
`rmarkdown::render("mcpower/validation/validation_get_effects.rmd", output_dir = "mcpower/documentation/validation")`.
