---
output:
  md_document:
    variant: gfm+tex_math_dollars
---

# MCPower — Validating the Data Generator

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
suppressPackageStartupMessages(library(lme4))
# Run with the working directory set to mcpower/validation/ (rmarkdown::render
# does this automatically from the .rmd's location).
source("common.R")    # build_model, content_hash, parsers, fit_sota, var_components
source("tolerances.R")# DGP_TOL — the central A<->B gates
source("formulas.R")  # CASES — the catalogue of formulas to validate

# --- How many times each formula is regenerated -------------------------------
K_FAST <- 1600L  # ordinary + logistic models (matches the OLS simulation default)
K_LME  <- 800L   # mixed models (matches the mixed-model default; lmer is slower)
# A<->B gates, sourced centrally from tolerances.R (one moment_abs covers every
# absolute moment check; var_rel for the variance; bias_fdr_q/bias_logit_abs for
# beta recovery — split by family, see recovery_decision/beta_rows below).
BANDS  <- list(pred_mean = DGP_TOL$moment_abs, pred_sd = DGP_TOL$moment_abs,
               corr = DGP_TOL$moment_abs, prop = DGP_TOL$moment_abs,
               var_rel = DGP_TOL$var_rel, icc = DGP_TOL$moment_abs,
               icc_marginal = DGP_TOL$icc_marginal_abs)

# --- Generate one formula's data K times; collect statistics + fitted estimates
simulate_case <- function(case) {
  fam <- case$family
  K   <- if (fam == "lme") K_LME else K_FAST
  m   <- build_model(case)
  m$.debug_n <- as.integer(case$n); m$.debug_n_sims <- 1L
  pc <- parse_correlation(case); pf <- parse_factor_props(case)

  m$.debug_seed <- as.numeric(case$seed + 1L)        # probe draw: learn columns
  d0    <- m$create_data(); cols <- d0$columns
  p     <- length(cols)
  col_idx <- which(grepl("^col_", cols))
  # A 2-level factor becomes a binary col_ column; keep only genuinely continuous
  # col_ columns (>2 distinct values) for the average/spread checks.
  cont    <- col_idx[vapply(col_idx, function(i) length(unique(d0$design[, i])) > 2L, logical(1))]
  dummy   <- which(grepl("^factor_dummy_", cols))

  betas <- matrix(NA_real_, K, p); ses <- matrix(NA_real_, K, p)
  cmean <- matrix(NA_real_, K, length(cont)); csd <- matrix(NA_real_, K, length(cont))
  corr  <- if (!is.null(pc)) numeric(K) else NULL
  props <- if (!is.null(pf)) matrix(NA_real_, K, length(pf)) else NULL
  rvar  <- if (fam != "logit") numeric(K) else NULL
  icc   <- if (!is.null(case$cluster)) numeric(K) else NULL
  micc  <- if (!is.null(case$cluster)) numeric(K) else NULL  # observed (marginal) ICC of raw outcome
  pmarg <- if (!is.null(case$cluster)) numeric(K) else NULL  # predicted marginal ICC (same-draw components)
  tb    <- true_beta_vector(case, cols)                      # true betas (for Var(Xbeta))

  for (k in seq_len(K)) {
    m$.debug_seed <- as.numeric(case$seed + k)        # fresh random noise each draw
    d <- m$create_data()
    for (i in seq_along(cont)) {
      x <- d$design[, cont[i]]; cmean[k, i] <- mean(x); csd[k, i] <- stats::sd(x)
    }
    if (!is.null(pc)) corr[k] <- stats::cor(d$design[, pc$cols[1]], d$design[, pc$cols[2]])
    if (!is.null(pf)) {
      dm <- d$design[, dummy, drop = FALSE]
      props[k, ] <- c(mean(rowSums(dm) == 0), colMeans(dm))
    }
    if (fam == "ols") {
      fr <- stats::lm.fit(d$design, d$outcome)
      rvar[k] <- sum(fr$residuals^2) / (length(d$outcome) - p)
    }
    if (!is.null(case$cluster)) {
      xb <- as.numeric(d$design %*% tb)
      # Conditional ICC tau^2/(tau^2+sigma^2) from TRUE-beta residuals: removing
      # the fixed part with the known betas (not a cluster-naive OLS fit) keeps the
      # between component from being deflated on factor designs whose dummies have
      # high-variance cluster means (see common.R var_components).
      v <- var_components(d$outcome - xb, d$cluster_ids)
      icc[k] <- v$icc; rvar[k] <- v$within
      # Marginal (observed) ICC: the raw outcome's between/within (fixed part left
      # IN) -> below the conditional ICC because Var(Xbeta) inflates the total.
      micc[k] <- var_components(d$outcome, d$cluster_ids)$icc
      # Predicted marginal from the SAME draw: compose the budget from this draw's
      # conditional components (between = tau^2, within = sigma^2) PLUS the fixed
      # part's OWN between/within split. Var(Xbeta) is NOT all within-cluster when
      # a factor's level mix varies across clusters, so its between part joins the
      # numerator; for continuous predictors between(Xbeta) ~ 0 and this reduces to
      # tau^2/(tau^2+sigma^2+Var(Xbeta)).
      vx       <- var_components(xb, d$cluster_ids)
      pmarg[k] <- (v$between + vx$between) /
                  (v$between + vx$between + v$within + vx$within)
    }
    f <- fit_sota(fam, d$design, d$outcome, d$cluster_ids)
    betas[k, ] <- f$beta; ses[k, ] <- f$se
  }
  list(case = case, K = K, cols = cols, terms = term_labels(case, cols),
       cont = cont, dummy = dummy, pc = pc, pf = pf,
       betas = betas, ses = ses, cmean = cmean, csd = csd,
       corr = corr, props = props, rvar = rvar, icc = icc,
       micc = micc, pmarg = pmarg, true_beta = tb)
}

# --- Plain-language renderings of a formula -----------------------------------
# The true model as an equation, e.g. "y = 0.25·x1 + 0.10·x2 + noise".
model_equation <- function(case) {
  e <- parse_effects(case$effects)
  parts <- vapply(seq_along(e), function(i)
    sprintf("%s %g·%s", if (e[i] < 0) "−" else "+", abs(e[i]), names(e)[i]), character(1))
  body <- paste(parts, collapse = " ")
  if (case$family == "logit") {
    sprintf("log-odds(y = 1) = logit(%.2f) %s", case$baseline_probability, body)
  } else if (case$family == "lme") {
    sprintf("y = %s + per-%s random intercept (ICC %.2f) + noise",
            sub("^\\+ ", "", body), case$cluster$var, case$cluster$ICC)
  } else {
    sprintf("y = %s + noise", sub("^\\+ ", "", body))
  }
}

family_long <- function(f) switch(f,
  ols = "ordinary least squares (R's `lm`)",
  logit = "logistic regression (R's `glm`)",
  lme = "linear mixed model (`lme4::lmer`)")

factor_props_of <- function(case, f) {
  decls <- unlist(strsplit(case$variable_types, ";"))
  d <- decls[grepl(paste0("^\\s*", f, "\\s*="), decls)][1]
  mm <- regmatches(d, regexpr("factor,([^)]*)", d))
  as.numeric(trimws(strsplit(sub("factor,", "", mm), ",")[[1]]))
}

build_description <- function(case) {
  cont <- continuous_names(case); fac <- factor_names(case); pieces <- character(0)
  if (length(cont))
    pieces <- c(pieces, sprintf("%s (%s) drawn from a standard normal (mean 0, sd 1)",
      if (length(cont) == 1) "one continuous predictor" else paste(length(cont), "continuous predictors"),
      paste0("`", cont, "`", collapse = ", ")))
  if (!is.null(case$correlations)) {
    pc <- parse_correlation(case)
    pieces <- c(pieces, sprintf("`%s` and `%s` correlated at %.2f",
      pc$vars[1], pc$vars[2], pc$rho))
  }
  for (f in fac) {
    pr <- factor_props_of(case, f)
    pieces <- c(pieces, sprintf("a %d-level factor `%s` at proportions %s",
      length(pr), f, paste0(round(100 * pr), "%", collapse = "/")))
  }
  outcome <- switch(case$family,
    ols  = "The outcome is the formula above plus standard-normal noise.",
    logit= sprintf("The outcome is 1 or 0, drawn at the logistic probability of the formula above (baseline rate %d%%).",
                   round(100 * case$baseline_probability)),
    lme  = sprintf(paste0("The outcome is the formula above, plus a shared offset for each of the %d clusters ",
                   "(%d observations each, sized to an intra-cluster correlation of %.2f), plus standard-normal noise."),
                   case$cluster$n_clusters, case$cluster$cluster_size, case$cluster$ICC))
  sprintf("**How it's built:** %s. %s Sample size %d, random seed %d.",
          paste(pieces, collapse = "; "), outcome, case$n, case$seed)
}

# --- Per-formula result tables ------------------------------------------------
desc_rows <- function(s) {
  rows <- list()
  add <- function(property, requested, vals, band, kind) {
    avg <- mean(vals); off <- abs(avg - requested)
    ok  <- if (kind == "rel") off <= band * abs(requested) else off <= band
    rows[[length(rows) + 1L]] <<- data.frame(
      `What we check`      = property,
      `Requested`          = requested,
      `Average over draws` = avg,
      `Allowed difference` = if (kind == "rel") sprintf("within %g%%", band * 100) else sprintf("within %g", band),
      `Result`             = if (ok) "OK" else "OFF",
      check.names = FALSE)
  }
  for (i in seq_along(s$cont)) {
    add(sprintf("average of %s",      s$terms[s$cont[i]]), 0, s$cmean[, i], BANDS$pred_mean, "abs")
    add(sprintf("std. deviation of %s", s$terms[s$cont[i]]), 1, s$csd[, i], BANDS$pred_sd, "abs")
  }
  if (!is.null(s$corr))
    add(sprintf("correlation of %s", paste(s$pc$vars, collapse = " and ")), s$pc$rho, s$corr, BANDS$corr, "abs")
  if (!is.null(s$pf))
    for (j in seq_along(s$pf))
      add(sprintf("proportion in level %d", j - 1L), s$pf[j], s$props[, j], BANDS$prop, "abs")
  if (!is.null(s$rvar))
    add(if (!is.null(s$case$cluster)) "within-cluster variance" else "noise variance", 1, s$rvar, BANDS$var_rel, "rel")
  if (!is.null(s$icc))
    add("intra-cluster correlation (ICC)", s$case$cluster$ICC, s$icc, BANDS$icc, "abs")
  if (!is.null(s$micc))
    # The set ICC is the CONDITIONAL ICC tau^2/(tau^2+sigma^2); the raw outcome's
    # (marginal) ICC is lower because the fixed part adds Var(Xbeta) to the total.
    # s$pmarg is the per-draw prediction composing tau^2, sigma^2 and the fixed
    # part's between/within split (simulate_case); we check the K-draw mean of the
    # observed marginal lands on it -- i.e. the variance budget composes.
    add("observed (marginal) ICC vs predicted", mean(s$pmarg), s$micc, BANDS$icc_marginal, "abs")
  do.call(rbind, rows)
}

# Global A<->B recovery decision. OLS/LME: pool every coefficient's two-sided
# z-of-the-mean p-value across ALL such cases and control the FDR via
# Benjamini-Hochberg at DGP_TOL$bias_fdr_q (a single hard per-coef z-gate over
# ~103 coefficients would abort ~all clean renders on noise). logit: absolute band
# (its finite-sample bias is systematic, not a detectable-vs-null effect).
recovery_decision <- function(RES) {
  rows <- do.call(rbind, lapply(RES, function(s) {
    mb <- colMeans(s$betas); sb <- apply(s$betas, 2, stats::sd)
    z  <- (mb - s$true_beta) / (sb / sqrt(s$K))
    data.frame(key = paste(s$case$label, s$terms, sep = "::"),
               family = s$case$family, z = z,
               absdiff = abs(mb - s$true_beta),
               stringsAsFactors = FALSE)
  }))
  off <- logical(nrow(rows))
  is_logit <- rows$family == "logit"
  off[is_logit]  <- rows$absdiff[is_logit] > DGP_TOL$bias_logit_abs
  p <- 2 * stats::pnorm(-abs(rows$z[!is_logit]))            # two-sided; ~Uniform under clean DGP
  off[!is_logit] <- stats::p.adjust(p, method = "BH") <= DGP_TOL$bias_fdr_q
  setNames(off, rows$key)
}

beta_rows <- function(s, off = RECOVERY_OFF) {
  mb <- colMeans(s$betas); sb <- apply(s$betas, 2, stats::sd)
  z  <- (mb - s$true_beta) / (sb / sqrt(s$K))
  keys   <- paste(s$case$label, s$terms, sep = "::")
  result <- ifelse(off[keys], "OFF", "OK")
  data.frame(
    `Term`                  = s$terms,
    `True value`            = s$true_beta,
    `Recovered (average)`   = mb,
    `Spread across draws`   = sb,
    `Std. errors from true` = z,
    `Result`                = result,
    check.names = FALSE, row.names = NULL)
}

recovery_note <- function(s) {
  br <- beta_rows(s); eff <- br[br$Term != "intercept", ]
  bad <- eff[eff$Result != "OK", ]
  msg <- if (nrow(bad)) {
    if (s$case$family == "logit")
      sprintf(paste0("A few coefficients sit slightly off the true value (%s). This is the small, well-known ",
              "finite-sample bias of logistic-regression estimates — expected, not a generator fault."),
              paste(bad$Term, collapse = ", "))
    else
      sprintf(paste0("Some coefficients were flagged by the pooled Benjamini-Hochberg recovery gate (%s) — ",
              "worth investigating."),
              paste(bad$Term, collapse = ", "))
  } else {
    "Every coefficient is centred on its true value — the formula holds in the generated data."
  }
  if (!is.null(s$pc))
    msg <- c(msg, sprintf(paste0("Because `%s` and `%s` are correlated (%.2f), their estimates vary more from ",
            "draw to draw; that wider spread is the expected effect of collinearity, not a problem."),
            s$pc$vars[1], s$pc$vars[2], s$pc$rho))
  paste(msg, collapse = " ")
}

# Conditional-vs-marginal ICC explainer, emitted for cluster cases only (next to
# the two ICC rows). Frames the gap as expected, like the logistic-bias note.
icc_note <- function(s) {
  if (is.null(s$micc)) return(NULL)
  paste0(
    "The intra-cluster correlation you set is the **conditional** ICC — the ",
    "correlation between two observations in the same cluster *after accounting ",
    "for the predictors* — and the generator recovers it (the \"intra-cluster ",
    "correlation (ICC)\" row above). The **observed (marginal) ICC** of the raw ",
    "outcome is lower, because the predictors explain part of the total variance; ",
    "the stronger the fixed effects, the larger the gap. This is the standard ",
    "conditional-vs-marginal distinction — expected and correct, not a generation ",
    "fault — and the \"observed (marginal) ICC vs predicted\" row confirms the ",
    "observed value lands on what that distinction predicts.")
}

plot_recovery <- function(s) {
  eff <- which(s$cols != "intercept")
  op  <- graphics::par(mfrow = c(1, length(eff)), mar = c(4, 4, 2.5, 1)); on.exit(graphics::par(op))
  for (j in eff) {
    b <- s$betas[, j]
    graphics::hist(b, breaks = 40, col = "grey85", border = "white",
                   main = s$terms[j], xlab = "recovered coefficient")
    graphics::abline(v = s$true_beta[j], col = "red",  lwd = 2)
    graphics::abline(v = mean(b),        col = "blue", lwd = 2, lty = 2)
  }
}

# Regenerating the saved dataset from its seed must reproduce its fingerprint.
hash_check <- function(case) {
  saved <- readRDS(file.path("data", paste0(case$label, ".rds")))
  m <- build_model(case)
  m$.debug_n <- as.integer(case$n); m$.debug_n_sims <- 1L; m$.debug_seed <- as.numeric(case$seed)
  d <- m$create_data()
  list(ok = identical(content_hash(d$design, d$outcome, d$cluster_ids), saved$hash))
}
```

# What this report shows

MCPower answers power-analysis questions by **generating synthetic datasets** from
a formula you specify — for example *"the outcome `y` equals `0.25 × x1` plus
random noise"* — and then simulating your analysis on those datasets many times.
Everything downstream depends on one thing being true: **the data it generates
must actually embody the formula you asked for.** If the generator is even
slightly wrong, every power number built on top of it is wrong too.

This report checks exactly that. For each supported formula we *chose* the true
coefficients, so we know the right answer in advance. The question is simple:

> When an independent, standard statistical analysis is run on MCPower's
> generated data, does it recover the true numbers we put in?

If it does — consistently, across every supported kind of model — the generator
is faithful.

## How the data is generated

For each formula MCPower builds a dataset of `n` rows:

- **continuous predictors** are drawn from a standard normal distribution
  (average 0, standard deviation 1);
- **categorical predictors (factors)** are assigned at the requested proportions
  (a 2-level factor becomes a single 0/1 column; a 3-level factor becomes two);
- **clustered (mixed-effects) data** groups the rows and gives each group a shared
  random offset, sized to hit the requested **intra-cluster correlation (ICC)** —
  specifically the *conditional* ICC, the correlation that remains after the
  predictors are accounted for (the raw outcome's *marginal* ICC is lower whenever
  the predictors explain variance);
- the **outcome** is computed from the true formula and coefficients, then random
  variation is added — standard-normal noise for ordinary regression, a 0/1 draw
  at the model's probability for logistic regression, a cluster offset plus noise
  for mixed models.

A single dataset is noisy, so we never judge from one. Instead each formula is
**regenerated many times with fresh random noise** — `r format(K_FAST, big.mark=",")`
times for ordinary and logistic models, `r format(K_LME, big.mark=",")` for mixed
models — and we look at the *distribution* of results across all of those draws.

## The three checks

**1. Data checks.** The basic statistics of the generated data should match what
was requested: predictor averages near 0, standard deviations near 1, the
requested correlations and factor proportions, the noise variance, and the ICC.
We **average each statistic over all draws** (which cancels the random scatter of
any single dataset) and require it to land within a small fixed tolerance of the
requested value.

**2. Coefficient recovery** — the headline check: *is the formula actually true
in the data?* For every generated dataset we fit a completely **independent,
standard model in R** (`lm`, `glm`, or `lme4::lmer`) and read off the estimated
coefficients. Averaged over all draws, each estimate should land on the true
value we put in. We measure how far the average estimate is from the truth in
units of its own **standard error** (the estimate's draw-to-draw scatter divided
by the square root of the number of draws). For OLS and mixed-effects models, the
~103 per-coefficient z-scores are pooled across all cases and the
**false-discovery rate** is controlled via Benjamini-Hochberg at q = `r DGP_TOL$bias_fdr_q`
(a fixed per-coefficient z-threshold below ~4 would abort nearly every clean
render on sampling noise alone). For logistic regression we use an **absolute
band** instead (|mean − true| ≤ `r DGP_TOL$bias_logit_abs`), because logistic
MLE's finite-sample bias swamps the z-score at this K — see below.

**3. File reproduces.** The dataset saved to disk for each formula is regenerated
from its random seed, and its content fingerprint (a SHA-256 hash) is compared to
the saved one. A mismatch would mean the generator's output had drifted, so any
later comparison against the saved file would be meaningless.

## The thresholds

| What we check | Requested value | Allowed difference |
|---|---|---|
| average of each continuous predictor | 0 | within `r BANDS$pred_mean` |
| standard deviation of each predictor | 1 | within `r BANDS$pred_sd` |
| correlation between predictors | the value you set | within `r BANDS$corr` |
| factor level proportions | the values you set | within `r BANDS$prop` |
| noise / within-cluster variance | 1 | within `r BANDS$var_rel * 100`% |
| intra-cluster correlation (ICC), conditional | the value you set | within `r BANDS$icc` |
| observed (marginal) ICC vs predicted | τ²/(τ²+σ²+Var(Xβ)) | within `r BANDS$icc_marginal` |
| **coefficient recovery (OLS/LME)** | the true coefficient | pooled BH-FDR ≤ **`r DGP_TOL$bias_fdr_q`** across all cases |
| **coefficient recovery (logit)** | the true coefficient | absolute difference within **`r DGP_TOL$bias_logit_abs`** |

The data-check tolerances are tight because they apply to the *average over all
draws*, not to a single dataset — averaging thousands of draws removes the random
scatter, so anything outside these bands signals a real generation problem.

**A note on logistic regression.** Logistic-regression estimates carry a small,
well-known *finite-sample bias*: with a limited number of rows the
maximum-likelihood estimate is slightly off even when everything is correct. With
this many draws that tiny bias becomes statistically visible (the logistic models
below show recovery distances running a bit higher, around 3–4 standard errors),
but it stays within tolerance at the sample sizes used here. It is an expected
statistical artefact, not a flaw in the generator.

**Bias vs. spread.** For each formula we report two different things separately.
*Bias* is whether the **average** estimate is off the true value (a centring
problem). *Spread* is how much the estimates **vary** from draw to draw. When two
predictors are correlated, their estimates naturally vary more — collinearity
inflates the spread. That wider spread is expected and correct; it is not a
failure, and we say so wherever it appears.

```{r compute}
RES  <- stats::setNames(lapply(CASES, simulate_case), names(CASES))
HASH <- stats::setNames(lapply(CASES, hash_check),    names(CASES))
RECOVERY_OFF <- recovery_decision(RES)
```

# Results at a glance

Every row is one formula with one set of true coefficients. "Data checks" and
"Coefficient recovery" summarise the per-formula detail that follows.

```{r summary}
summ <- data.frame(
  `Formula`              = vapply(RES, function(s) sprintf("`%s`", s$case$formula), character(1)),
  `True coefficients`    = vapply(RES, function(s) sprintf("`%s`", s$case$effects), character(1)),
  `n`                    = vapply(RES, function(s) as.integer(s$case$n), integer(1)),
  `Data checks`          = vapply(RES, function(s) if (all(desc_rows(s)$Result == "OK")) "all OK" else "review", character(1)),
  `Coefficient recovery` = vapply(RES, function(s) if (all(beta_rows(s)$Result == "OK")) "all OK" else "review", character(1)),
  `File reproduces`      = vapply(names(RES), function(l) if (HASH[[l]]$ok) "yes" else "NO", character(1)),
  row.names = NULL, check.names = FALSE)
knitr::kable(summ)
```

```{r overall, results='asis'}
all_ok <- all(vapply(RES, function(s) all(desc_rows(s)$Result == "OK") && all(beta_rows(s)$Result == "OK"), logical(1))) &&
          all(vapply(HASH, function(h) h$ok, logical(1)))
cat(if (all_ok)
  sprintf(paste0("> **All %d formulas pass all three checks.** The generated data matches the requested ",
          "statistics, an independent R model recovers every true coefficient, and every saved dataset ",
          "reproduces from its seed.\n"), length(RES))
  else "> **Some formulas need review** — see the per-formula detail below.\n")
stopifnot(all_ok)   # G5
```

# Formula-by-formula detail

```{r per-formula, results='asis'}
# One asis section per formula: heading is the true equation, then the build
# recipe, the data checks, the coefficient-recovery table + plot, and the file
# fingerprint check. The recovery plots are written into the published docs tree
# (../documentation/validation/figures/, beside this report's HTML) and referenced
# by a relative "figures/<lbl>.png" src — which resolves to /validation/figures/...
# at the served leyline page (UGCPolicy drops data: URIs, so they can't be inlined).
figdir_src <- "figures"
figdir_out <- file.path("..", "documentation", "validation", figdir_src)
dir.create(figdir_out, recursive = TRUE, showWarnings = FALSE)
png_type <- if (capabilities("cairo")) "cairo" else "Xlib"

for (lbl in names(RES)) {
  s <- RES[[lbl]]; case <- s$case; h <- HASH[[lbl]]
  cat(sprintf("\n## %s\n\n", model_equation(case)))
  cat(sprintf("R formula `%s` &middot; %s &middot; n = %d &middot; saved as `data/%s.rds`\n\n",
              case$formula, family_long(case$family), case$n, lbl))
  cat(build_description(case), "\n\n")

  cat("**Data checks** — statistics of the generated data, averaged over all draws, vs. what was requested:\n\n")
  cat(knitr::kable(desc_rows(s), digits = 4, format = "pipe"), sep = "\n"); cat("\n\n")

  cat("**Coefficient recovery** — an independent R model fitted to each generated dataset; the average estimate should land on the true value:\n\n")
  cat(knitr::kable(beta_rows(s), digits = 4, format = "pipe"), sep = "\n"); cat("\n\n")

  cat(recovery_note(s), "\n\n")
  icn <- icc_note(s); if (!is.null(icn)) cat(icn, "\n\n")

  fw <- max(4, 2.6 * sum(s$cols != "intercept"))
  fpath <- file.path(figdir_out, paste0(lbl, ".png"))
  grDevices::png(fpath, width = fw * 110, height = 320, res = 96, type = png_type)
  plot_recovery(s)
  grDevices::dev.off()
  cat(sprintf("![Recovered coefficients for %s over %d datasets.](%s)\n\n",
              case$formula, s$K, file.path(figdir_src, paste0(lbl, ".png"))))
  cat(sprintf("*Each panel: the spread of one recovered coefficient across %d generated datasets. Red line = the true value, blue dashed = the average estimate. File reproduces from seed: **%s**.*\n\n",
              s$K, if (h$ok) "yes" else "NO"))
}
```

# How this was produced

```{r provenance}
prov <- data.frame(
  item  = c("Report generated", "R version", "mcpower", "lme4",
            "Draws per formula (ordinary / logistic)", "Draws per formula (mixed)",
            "Recovery threshold", "Formulas validated"),
  value = c(format(Sys.Date(), "%e %B %Y"), R.version.string,
            as.character(utils::packageVersion("mcpower")),
            as.character(utils::packageVersion("lme4")),
            format(K_FAST, big.mark = ","), format(K_LME, big.mark = ","),
            sprintf("OLS/LME: pooled BH-FDR ≤ %g (Benjamini-Hochberg); logit: |mean−true| ≤ %g (absolute)",
                    DGP_TOL$bias_fdr_q, DGP_TOL$bias_logit_abs), as.character(length(RES))),
  check.names = FALSE)
knitr::kable(prov)
```

The datasets are generated by `mcpower/validation/data_generation.r` from the
formula catalogue in `mcpower/validation/formulas.R`; this report regenerates
them many times over and runs the checks above. To reproduce it, from the
repository root:

```r
rmarkdown::render("mcpower/validation/validation_data_generation.rmd",
                  output_dir = "mcpower/documentation/validation")
```
