---
output:
  md_document:
    variant: gfm+tex_math_dollars
---

# MCPower — Validating Uploaded-Data Generation

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Run with the working directory set to mcpower/validation/ (rmarkdown::render
# does this automatically from the .rmd's location).
source("common.R")    # build_model, content_hash, parsers
source("tolerances.R")# DGP_TOL
source("formulas.R")  # UPLOAD_CASES

# Number of simulated draws per case — matches the OLS simulation default.
K <- 1600L

# Reference draw size for measuring engine population moments (oracle).
# Large enough that the sample moments converge to the engine's population
# moments to well within DGP_TOL$moment_abs.
N_REF <- 200000L

# A<->B gates from tolerances.R, applied to engine-population vs draw averages.
BANDS <- list(
  mean    = DGP_TOL$moment_abs,   # average generated mean vs oracle mean
  sd      = DGP_TOL$moment_abs,   # average generated sd vs oracle sd
  prop    = DGP_TOL$moment_abs,   # binary proportion vs oracle proportion
  corr_cc = DGP_TOL$moment_abs    # cont×cont generated correlation vs oracle
)

# Helpers moved to common.R: upload_csv_abs, load_upload_frame, frame_raw_moments,
# build_upload_model, engine_oracle, simulate_upload_case, upload_assertion_rows.
# BANDS is a local alias over DGP_TOL; define here so the rmd prose can inline it.
```

# What this report validates

MCPower can drive its data generator from an **uploaded empirical frame**.
When you call `upload_data(pilot_data)`, the engine:

1. **Standardizes** each continuous predictor to mean 0, sd 1 (population SD,
   ddof = 0), and stores the result as a sorted lookup table for NORTA quantile
   mapping.
2. **Measures** empirical Pearson correlations among **continuous** predictors
   and installs them as the NORTA latent-correlation matrix (under `mode = "partial"`,
   the default). Binary/factor predictors are generated from their marginals only —
   they are never correlated with any other predictor.
3. For each draw, maps correlated standard-normal variates through the
   **empirical inverse-CDF** of the stored (standardized) frame column — producing
   data that follows the frame's marginal distribution.

## A note on the oracle: engine moments ≠ raw frame moments

For a **small frame** (the 32-row mtcars pilot), the NORTA quantile path
introduces two well-known finite-frame effects:

- **Mean shift**: `E[Q̂(Φ(Z))]` for Z∼N(0,1) is not exactly 0, because the
  type-7 empirical quantile of a finite sorted sample maps Uniform(0,1) input to
  a slightly off-centre discrete distribution.  For the 32-row mtcars frame,
  the mean of continuous columns is shifted by ≈−0.025 — well outside the raw
  frame's standardized mean of 0.

- **Variance shrinkage**: The empirical quantile is bounded by the frame's
  observed min and max. The tails of Φ(Z) map to the extremes of the 32-point
  sample rather than the continuous normal tails, so the generated standard
  deviation is ≈0.93 instead of 1.0.

- **Binary predictors are independent by design**: Correlation is continuous-only
  in MCPower's engine.  Binary/factor variables are generated from their marginals,
  independent of every other predictor.  The realized binary×continuous Pearson r
  is therefore expected to be ≈ 0, and this is numerically gated (see §Binary below).

**These are expected properties of the NORTA algorithm with a finite reference
frame, not engine bugs.** The design decision (NORTA + empirical inverse-CDF) is
what makes the generated distribution "look like" the frame distribution at large
N, even if the first two moments do not match the standardized frame exactly for
small frames.

Following the principle that MCPower is its own golden source, the validation oracle is
**what the engine itself produces at large N** (a single n = `r format(N_REF, big.mark=",")`
reference draw), not the raw frame statistics. The A↔B assertion is: the average
of K = `r K` draws at n = 400 must reproduce the engine's large-N moments within
`DGP_TOL$moment_abs = r DGP_TOL$moment_abs`.

The correlation checks in this report are engine-to-engine (large-N self-consistency); independent confirmation that NORTA installs the frame's empirical Spearman rank correlation (converted to the latent Gaussian scale) as the contract matrix, while binary predictors keep their correlation slots at zero, is an **L1 engine test** (`engine-spec-builder/tests/upload_test.rs::partial_mode_measures_nontrivial_and_excludes_binary`), not an L3 gate.

## What is and is not checked

| Assertion type | Checked? | Notes |
|---|---|---|
| Continuous marginal mean | **Yes** | K-draw average mean vs engine oracle |
| Continuous marginal sd | **Yes** | K-draw average sd vs engine oracle |
| Continuous × continuous correlation | **Yes** | K-draw average vs engine oracle |
| Binary marginal proportion | **Yes** | K-draw average proportion vs oracle |
| Binary × continuous correlation | **Yes (≈ 0 gate)** | Binary is independent by design; K-avg |r| must be ≤ `BANDS$corr_cc`; see §Binary |
| Factor correlations | **Out of scope** | Discrete-correlation work not yet landed |
| Nonlinear joint (parabola, strict mode) | **Yes** | `upload_strict_nonlinear`: strict bootstrap copies whole rows → generated rows lie on x2=x1² to 1e-5; NORTA would give O(1) residuals |

## Tolerance thresholds

Gates from `tolerances.R` (`DGP_TOL`), applied to the K-draw average vs the
engine oracle:

| What we check | Tolerance |
|---|---|
| Average generated mean of continuous column | within `r BANDS$mean` of oracle |
| Average generated sd of continuous column | within `r BANDS$sd` of oracle |
| Average generated binary proportion | within `r BANDS$prop` of oracle |
| Average generated cont×cont correlation | within `r BANDS$corr_cc` of oracle |
| Average generated binary×cont correlation | |K-avg r| ≤ `r BANDS$corr_cc` (oracle = 0; binary is independent) |

```{r compute, cache=FALSE}
# Run the upload simulation campaign.
RES <- stats::setNames(lapply(UPLOAD_CASES, simulate_upload_case), names(UPLOAD_CASES))
```

# Results at a glance

```{r summary}
summ <- data.frame(
  `Case`          = names(RES),
  `Formula`       = vapply(RES, function(s) sprintf("`%s`", s$case$formula), character(1)),
  `Upload cols`   = vapply(RES, function(s) paste(s$case$upload_cols, collapse = ", "), character(1)),
  `Assertions`    = vapply(RES, function(s) {
    ar <- upload_assertion_rows(s)
    if (all(ar$Verdict == "PASS")) sprintf("all %d PASS", nrow(ar)) else {
      n_fail <- sum(ar$Verdict == "FAIL")
      sprintf("%d FAIL / %d total", n_fail, nrow(ar))
    }
  }, character(1)),
  row.names = NULL, check.names = FALSE
)
knitr::kable(summ)
```

```{r overall, results='asis'}
all_pass <- all(vapply(RES, function(s) all(upload_assertion_rows(s)$Verdict == "PASS"), logical(1)))
cat(if (all_pass)
  sprintf(paste0("> **All %d upload cases pass all assertions.** ",
          "NORTA-over-frame (partial/none) and strict-bootstrap paths are ",
          "internally consistent: K = %d draws per case reproduce the engine's ",
          "own large-N moments within DGP_TOL, and strict-mode rows lie on the ",
          "uploaded parabola to within 1e-5.\n"), length(RES), K)
  else "> **Some assertions fail — see per-case detail below.**\n")
```

# Case-by-case detail

```{r per-case, results='asis'}
for (lbl in names(RES)) {
  s    <- RES[[lbl]]
  case <- s$case
  or   <- s$oracle
  raw  <- s$raw_fm

  cat(sprintf("\n## %s\n\n", case$label))
  cat(sprintf("Formula `%s` | OLS | upload columns: %s | n = %d per draw | K = %d draws\n\n",
              case$formula, paste(case$upload_cols, collapse = ", "), case$n, K))

  # --- Frame vs oracle moments comparison ---
  cat("### Frame moments vs engine oracle\n\n")
  cat(paste0(
    "The table below compares the **raw frame** statistics (before any engine processing) ",
    "to the **engine oracle** (moments measured from a single large n = ",
    format(N_REF, big.mark = ","), " draw). ",
    "Differences arise from the finite-frame NORTA effects described in the introduction.\n\n"
  ))

  comp_rows <- do.call(rbind, lapply(case$cont_cols, function(nm) {
    pop_mean_std <- 0.0   # standardized frame: exact mean=0
    pop_sd_std   <- 1.0   # standardized frame: exact sd=1 (pop sd)
    data.frame(
      Column = nm, Statistic = c("mean", "sd"),
      `Frame (standardized)` = c(pop_mean_std, pop_sd_std),
      `Engine oracle`        = c(or$oracle_mean[nm], or$oracle_sd[nm]),
      `Deviation`            = c(abs(or$oracle_mean[nm] - pop_mean_std),
                                  abs(or$oracle_sd[nm]   - pop_sd_std)),
      check.names = FALSE, stringsAsFactors = FALSE)
  }))
  if (!is.null(raw$bin_prop) && length(raw$bin_prop) > 0L) {
    for (nm in case$binary_cols) {
      comp_rows <- rbind(comp_rows, data.frame(
        Column = nm, Statistic = "proportion",
        `Frame (standardized)` = raw$bin_prop[nm],
        `Engine oracle`        = or$oracle_prop[nm],
        `Deviation`            = abs(or$oracle_prop[nm] - raw$bin_prop[nm]),
        check.names = FALSE, stringsAsFactors = FALSE))
    }
  }
  if (!is.na(or$oracle_corr_cc)) {
    pair_nm <- paste(case$cont_cols[1], "x", case$cont_cols[2])
    frame_r <- raw$cont_corr[1]
    comp_rows <- rbind(comp_rows, data.frame(
      Column = pair_nm, Statistic = "Pearson r",
      `Frame (standardized)` = frame_r,
      `Engine oracle`        = or$oracle_corr_cc,
      `Deviation`            = abs(or$oracle_corr_cc - frame_r),
      check.names = FALSE, stringsAsFactors = FALSE))
  }
  cat(knitr::kable(comp_rows, digits = 4, format = "pipe"), sep = "\n"); cat("\n\n")

  # --- Assertion table (K-draw avg vs oracle) ---
  cat("### Assertion table — K-draw average vs engine oracle\n\n")
  cat(sprintf("Averaged over %d draws of n = %d:\n\n", K, case$n))
  ar <- upload_assertion_rows(s)
  cat(knitr::kable(ar, digits = 4, format = "pipe"), sep = "\n"); cat("\n\n")

  # --- Nonlinear parabola-preservation section (strict-mode only) ---
  if (!is.null(s$parabola_max_resid)) {
    pc <- s$para_coef
    cat("### Parabola-preservation assertion — strict bootstrap preserves the nonlinear joint\n\n")
    cat(paste0(
      "This case uses `mode = \"strict\"` (whole-row bootstrap). Because the engine copies ",
      "entire rows from the uploaded frame, every generated row satisfies x2 = x1^2 exactly — ",
      "the nonlinear joint is preserved.\n\n",
      "In **standardized** coordinates (mean 0, pop-sd 1), the parabola becomes:\n\n",
      "    x2_std = ", sprintf("%.6f", pc$a), " * x1_std² + ",
                              sprintf("%.6f", pc$b), " * x1_std + ",
                              sprintf("%.6f", pc$c0), "\n\n",
      "We check `max |x2_std - f(x1_std)|` for every row in every draw and assert ",
      "that the mean of this per-draw maximum is ≤ 1e-5 (the CSV floating-point floor).\n\n",
      "**Why this discriminates strict from NORTA:** under `mode = \"partial\"`, the engine ",
      "generates rows via NORTA — it matches marginals and Pearson correlation but draws from ",
      "a Gaussian copula, so points are scattered around the parabola, not constrained to it. ",
      "The NORTA per-row residual from the parabola would be O(1) in standardized units; ",
      "the strict residual is ≤ 3e-8 (CSV float precision).\n\n"
    ))
    nb_rows <- data.frame(
      Metric        = "mean max |x2_std - f(x1_std)| over K draws",
      `K-draw value`= mean(s$parabola_max_resid),
      `Max observed`= max(s$parabola_max_resid),
      `Tolerance`   = DGP_TOL$parabola_abs,
      Verdict       = if (mean(s$parabola_max_resid) <= DGP_TOL$parabola_abs) "PASS" else "FAIL",
      check.names = FALSE, stringsAsFactors = FALSE)
    cat(knitr::kable(nb_rows, digits = 10, format = "pipe"), sep = "\n"); cat("\n\n")
  }

  # --- Binary correlations (gated ≈ 0) --- {#binary}
  if (!is.null(s$corr_bc) && !is.null(s$bc_pairs)) {
    cat("### Binary correlations — gated to ≈ 0 {#binary}\n\n")
    cat(paste0(
      "Binary/factor predictors are generated from their marginals, independent of ",
      "every other predictor (correlation is continuous-only by design). ",
      "The realized binary×continuous Pearson r must therefore be ≈ 0. ",
      "Each pair below is asserted: |K-draw average r| ≤ `BANDS$corr_cc` (= ",
      BANDS$corr_cc, "). ",
      "These assertions are also included in the assertion table above.\n\n"
    ))
    bc_rows <- lapply(seq_len(nrow(s$bc_pairs)), function(r) {
      bi <- s$bc_pairs$b[r]; ci <- s$bc_pairs$c[r]
      bin_nm  <- case$binary_cols[bi]
      cont_nm <- case$cont_cols[ci]
      k_avg   <- mean(s$corr_bc[[r]])
      k_sd    <- stats::sd(s$corr_bc[[r]])
      ok      <- abs(k_avg) <= BANDS$corr_cc
      data.frame(
        Pair                    = sprintf("%s × %s", bin_nm, cont_nm),
        `K-draw mean r`         = k_avg,
        `K-draw sd r`           = k_sd,
        `Tolerance (|r| ≤)`     = BANDS$corr_cc,
        Verdict                 = if (ok) "PASS" else "FAIL",
        check.names = FALSE, stringsAsFactors = FALSE)
    })
    bc_tbl <- do.call(rbind, bc_rows)
    cat(knitr::kable(bc_tbl, digits = 4, format = "pipe"), sep = "\n"); cat("\n\n")
  }

  # --- mode="none" sanity check ---
  cat("### mode = \"none\" sanity check — marginals only, no correlation\n\n")
  cat(paste0(
    "Under `mode = \"none\"` the engine reproduces each predictor's marginal ",
    "independently (no correlation structure is imposed). We verify that continuous ",
    "column means and sds reproduce the engine's **none-mode** population moments — ",
    "measured from a matched large-n `mode = \"none\"` reference draw, *not* the ",
    "strict-mode oracle. The strict (whole-row bootstrap) and none (NORTA empirical-",
    "quantile) paths produce slightly different marginal distributions from the same ",
    "frame — the NORTA path carries a small (~0.5–1%) variance shrinkage the bootstrap ",
    "does not — so the none-mode draws must be gated against a none-mode oracle.\n\n"
  ))

  K_NONE  <- 400L
  frame_n <- load_upload_frame(case)
  m_none  <- build_upload_model(case, frame_n, mode = "none")
  m_none$.debug_n <- as.integer(case$n)
  n_cont <- length(case$cont_cols)

  # Oracle for the none-mode check must itself be measured under mode = "none".
  # `or` (= s$oracle) is the strict/partial-mode oracle from case$upload_mode; its
  # large-n SD is the frame's own (standardized) SD. The mode = "none" NORTA path
  # maps through the finite empirical quantile table and carries a small variance
  # shrinkage relative to it, so gating none-mode draws against the strict oracle
  # is apples-to-oranges. Draw a matched large-n none-mode reference so like is
  # compared with like (mirrors engine_oracle's N_REF draw + seed offset).
  m_none_oracle <- build_upload_model(case, frame_n, mode = "none")
  m_none_oracle$.debug_n    <- N_REF
  m_none_oracle$.debug_seed <- as.numeric(case$seed + 99999L)
  des_none <- m_none_oracle$create_data()$design
  oracle_none_mean <- vapply(or$cont_idx, function(j) mean(des_none[, j]), numeric(1))
  oracle_none_sd   <- vapply(or$cont_idx, function(j) stats::sd(des_none[, j]), numeric(1))

  cm_none <- matrix(NA_real_, K_NONE, n_cont)
  cs_none <- matrix(NA_real_, K_NONE, n_cont)
  for (k in seq_len(K_NONE)) {
    m_none$.debug_seed <- as.numeric(case$seed + k + K)
    d <- m_none$create_data()
    for (i in seq_len(n_cont)) {
      x <- d$design[, or$cont_idx[i]]
      cm_none[k, i] <- mean(x)
      cs_none[k, i] <- stats::sd(x)
    }
  }

  none_rows <- do.call(rbind, lapply(seq_len(n_cont), function(i) {
    nm <- case$cont_cols[i]
    m_ok <- abs(mean(cm_none[, i]) - oracle_none_mean[i]) <= BANDS$mean
    s_ok <- abs(mean(cs_none[, i]) - oracle_none_sd[i])   <= BANDS$sd
    rbind(
      data.frame(Column = nm, Stat = "mean",
                 Oracle = oracle_none_mean[i],
                 `K-avg` = mean(cm_none[, i]),
                 Diff = abs(mean(cm_none[, i]) - oracle_none_mean[i]),
                 Verdict = if (m_ok) "PASS" else "FAIL",
                 check.names = FALSE, stringsAsFactors = FALSE),
      data.frame(Column = nm, Stat = "sd",
                 Oracle = oracle_none_sd[i],
                 `K-avg` = mean(cs_none[, i]),
                 Diff = abs(mean(cs_none[, i]) - oracle_none_sd[i]),
                 Verdict = if (s_ok) "PASS" else "FAIL",
                 check.names = FALSE, stringsAsFactors = FALSE)
    )
  }))
  cat(knitr::kable(none_rows, digits = 4, format = "pipe"), sep = "\n"); cat("\n\n")
}
```

# Provenance

```{r provenance}
prov <- data.frame(
  Item  = c("Report generated", "R version", "mcpower",
            "Upload frames", "Oracle draw size",
            "Draws per case (mode=partial/strict)",
            "Draws per case (mode=none sanity)",
            "Moment tolerance",
            "Parabola tolerance (strict case)"),
  Value = c(format(Sys.Date(), "%e %B %Y"), R.version.string,
            as.character(utils::packageVersion("mcpower")),
            paste0("mtcars (built-in R dataset; 32 rows, hp/wt/am columns); ",
                   "nonlinear_parabola.csv (synthetic; 150 rows, x1∈[0,3], x2=x1²)"),
            format(N_REF, big.mark = ","),
            format(K, big.mark = ","),
            "400",
            sprintf("%.2g", DGP_TOL$moment_abs),
            "1e-5"),
  check.names = FALSE
)
knitr::kable(prov)
```

To reproduce this report, from the repository root:

```r
rmarkdown::render("mcpower/validation/validation_upload.rmd",
                  output_dir = "mcpower/web/documentation/validation")
```
